---
title: "GenePrediction"
author: "Jiabei Yang"
date: "5/23/2017"
output: pdf_document
---

```{r}
library(dplyr)
library(tidyr)
library(seqRFLP)
```


# 1 Separate the elites from the genetic information file

```{r}
gene_original_data <- read.csv("../Original/Genetic Information.csv")
prediction_data <- read.csv("../Elite/Prediction_StageII_best.csv")
elites_data_cls <- gene_original_data %>% filter(!is.na(CLASS_OF))
elites_data_noncls <- gene_original_data %>% filter(is.na(CLASS_OF))
elites_data_predict <- gene_original_data %>% 
  filter(VARIETY %in% prediction_data$VARIETY_ID[prediction_data$PREDICTION == 1])

elites_yes <- read.csv("../Elite/RMgroupS3_true13.csv")
elites_yes <- gene_original_data %>% 
  filter(VARIETY %in% elites_yes$VARIETY_ID[elites_yes$PREDICTION == 1])
varieties_no <- setdiff(gene_original_data$VARIETY, elites_yes$VARIETY)
elites_no <- gene_original_data %>% filter(VARIETY %in% varieties_no)
```

# 2 Output the elite data into FASTA format
```{r}
# cuttingLines_tmp1 <- seq(5, 2167, 40)
# cuttingLines_tmp2 <- seq(44, 2167, 40)
# cuttingLines <- NULL
# for (i in 1:length(cuttingLines_tmp2)){
#   cuttingLines <- c(cuttingLines, cuttingLines_tmp1[i], cuttingLines_tmp2[i])
# }
# cuttingLines <- c(cuttingLines, cuttingLines_tmp1[length(cuttingLines_tmp1)], 2167)
# 
# lineNames <- paste("Line", 1:(length(cuttingLines)/2), sep = "")
# for (i in 1:length(lineNames)){
#   elites_data_predict <- cbind(elites_data_predict, apply(elites_data_predict[, (cuttingLines[2*i-1]):(cuttingLines[2*i])], 1, paste, collapse = ""))
#   colnames(elites_data_predict)[dim(elites_data_predict)[2]] <- lineNames[i]
# }
# elites_data_predict <- elites_data_predict[, c("VARIETY", lineNames)]

elites_data_predict <- elites_data_predict %>% 
  mutate(sequence = apply(elites_data_predict[, 5:2167], 1, paste, collapse = ""))
elites_data_predict <- elites_data_predict[, c("VARIETY", "sequence")]

dataframe2fas(elites_data_predict, file = "Elites_predict_StageII.fasta")
```

# 3 Output class data and non-class data to fasta

```{r}
# All Elites
elites_data_overall <- elites_data_cls %>%
  mutate(sequence = apply(elites_data_cls[, 5:2167], 1, paste, collapse = ""))
elites_data_overall <- elites_data_overall[, c("VARIETY", "sequence")]

dataframe2fas(elites_data_overall, file = "Elites_overall.fasta")

# NonElites
NonElites_data_overall <- elites_data_noncls %>%
  mutate(sequence = apply(elites_data_noncls[, 5:2167], 1, paste, collapse = ""))
NonElites_data_overall <- NonElites_data_overall[, c("VARIETY", "sequence")]
NonElites_data_overall1 <- NonElites_data_overall[1:461, ]
NonElites_data_overall2 <- NonElites_data_overall[462:922, ]
NonElites_data_overall3 <- NonElites_data_overall[923:1383, ]

dataframe2fas(NonElites_data_overall, file = "NonElites_overall.fasta")
dataframe2fas(NonElites_data_overall1, file = "NonElites_overall1.fasta")
dataframe2fas(NonElites_data_overall2, file = "NonElites_overall2.fasta")
dataframe2fas(NonElites_data_overall3, file = "NonElites_overall3.fasta")

# Elites 2011
elites_data_cls2011 <- elites_data_cls %>%
  filter(CLASS_OF == 2011) 
elites_data_cls2011 <- elites_data_cls2011 %>%
  mutate(sequence = apply(elites_data_cls2011[, 5:2167], 1, paste, collapse = ""))
elites_data_cls2011 <- elites_data_cls2011[, c("VARIETY", "sequence")]

dataframe2fas(elites_data_cls2011, file = "Elites_cls2011.fasta")

# Elites 2012
elites_data_cls2012 <- elites_data_cls %>%
  filter(CLASS_OF == 2012) 
elites_data_cls2012 <- elites_data_cls2012 %>%
  mutate(sequence = apply(elites_data_cls2012[, 5:2167], 1, paste, collapse = ""))
elites_data_cls2012 <- elites_data_cls2012[, c("VARIETY", "sequence")]

dataframe2fas(elites_data_cls2012, file = "Elites_cls2012.fasta")

# Elites 2013
elites_data_cls2013 <- elites_data_cls %>%
  filter(CLASS_OF == 2013) 
elites_data_cls2013 <- elites_data_cls2013 %>%
  mutate(sequence = apply(elites_data_cls2013[, 5:2167], 1, paste, collapse = ""))
elites_data_cls2013 <- elites_data_cls2013[, c("VARIETY", "sequence")]

dataframe2fas(elites_data_cls2013, file = "Elites_cls2013.fasta")

# Elites 2014
elites_data_cls2014 <- elites_data_cls %>%
  filter(CLASS_OF == 2014) 
elites_data_cls2014 <- elites_data_cls2014 %>%
  mutate(sequence = apply(elites_data_cls2014[, 5:2167], 1, paste, collapse = ""))
elites_data_cls2014 <- elites_data_cls2014[, c("VARIETY", "sequence")]

dataframe2fas(elites_data_cls2014, file = "Elites_cls2014.fasta")
```

# Find patterns in Elites

We assume "NN" means we don't know about this location, and it won't be any other possible values.

```{r}
FindPattern <- function(elites_data_cls, elites_data_noncls){
  OneLevelElites <- NULL
  TwoLevelElites <- NULL
  ThreeLevelElites <- NULL
  OneLevelNonElites <- NULL
  TwoLevelNonElites <- NULL
  ThreeLevelNonElites <- NULL
  j <- 1
  k <- 1
  jj <- 1
  for (i in 5:2167){
    
    if (length(unique(elites_data_cls[, i])) == 1){
      OneLevelElites <- cbind(OneLevelElites, as.character(elites_data_cls[1, i]))
      colnames(OneLevelElites)[j] <- colnames(elites_data_cls)[i]
      j <- j+1
    }
    else if (length(unique(elites_data_cls[, i])) == 2){
      TwoLevelElites <- cbind(TwoLevelElites, as.character(unique(elites_data_cls[, i])))
      colnames(TwoLevelElites)[jj] <- colnames(elites_data_cls)[i]
      jj <- jj+1
    }
    
    if (length(unique(elites_data_noncls[, i])) == 1){
      OneLevelNonElites <- cbind(OneLevelNonElites, as.character(elites_data_noncls[1, i]))
      colnames(OneLevelNonElites)[k] <- colnames(elites_data_noncls)[i]
      k <- k+1
    }
  
  }
  
  patternOneLevel <- OneLevelElites[, setdiff(colnames(OneLevelElites), 
                                              colnames(OneLevelNonElites))]
  probPattern_noncls <- NULL
  for (i in 1:length(patternOneLevel)){
    probPattern_noncls <- c(probPattern_noncls, 
                            mean(elites_data_noncls[, names(patternOneLevel)[i]] == patternOneLevel[i]))
  }
  names(probPattern_noncls) <- names(patternOneLevel)
  
  probPattern_noncls <- data.frame(location = names(probPattern_noncls),
                                   probElite = probPattern_noncls,
                                   pattern = as.character(patternOneLevel))
  probPattern_noncls <- probPattern_noncls %>% arrange(probElite)
  
  probPattern2_noncls <- NULL
  for (i in 1:dim(TwoLevelElites)[2]){
    probPattern2_noncls <- c(probPattern2_noncls, 
                            mean(elites_data_noncls[, colnames(TwoLevelElites)[i]] %in% TwoLevelElites[,i]))
  }
  names(probPattern2_noncls) <- colnames(TwoLevelElites)
  
  probPattern2_noncls <- data.frame(location = names(probPattern2_noncls),
                                   probElite = probPattern2_noncls,
                                   pattern1 = as.character(TwoLevelElites[1,]),
                                   pattern2 = as.character(TwoLevelElites[2,]))
  probPattern2_noncls <- probPattern2_noncls %>% arrange(probElite)
  return(probPattern_noncls)
}
```

With 0.76^17 prob we will identify a wrong elite.

Prove:
```{r}
probPattern_noncls <- probPattern_noncls[1:84,]
probPattern2_noncls <- probPattern2_noncls[1:121,]
ElitesinNonElites <- elites_data_noncls
for (i in 1:dim(probPattern_noncls)[1]){
  ElitesinNonElites <- ElitesinNonElites %>% filter(as.character(eval(parse(text = paste("ElitesinNonElites", as.character(probPattern_noncls$location[i]), sep = "$")))) == as.character(probPattern_noncls$pattern[i]))
}

for (i in 1:dim(probPattern2_noncls)[1]){
  ElitesinNonElites <- ElitesinNonElites %>% filter(as.character(eval(parse(text = paste("ElitesinNonElites", as.character(probPattern2_noncls$location[i]), sep = "$")))) %in% as.character(unlist(probPattern2_noncls[i, c("pattern1", "pattern2")])))
}

```

Test on our predicted elite data.

```{r}
probPattern_noncls <- FindPattern(elites_yes, elites_no)

ElitesinNonElites <- elites_no
for (i in 1:dim(probPattern_noncls)[1]){
  ElitesinNonElites <- ElitesinNonElites %>% filter(as.character(eval(parse(text = paste("ElitesinNonElites", as.character(probPattern_noncls$location[i]), sep = "$")))) == as.character(probPattern_noncls$pattern[i]))
  if (dim(ElitesinNonElites)[1] / dim(elites_no)[1] <= 0.05){
    break
  }
}
print(i)
probPattern_noncls <- probPattern_noncls[1:i,]
```

```{r}
probPattern_noncls_overall <- FindPattern(elites_data_cls, elites_data_noncls)
```


1. all of them are duplicates

# Output Figures

```{r}
SignificantLoc <- data.frame(location = colnames(gene_original_data)[5:2167])
SignificantLoc <- left_join(SignificantLoc, probPattern_noncls, by = "location")
SignificantLoc <- SignificantLoc %>% 
  mutate(probElite = 1 - probElite) %>%
  mutate(probElite = ifelse(is.na(probElite), 0, probElite))
write.csv(SignificantLoc, "SigGenesFigure.csv", row.names = F)
```

